Structure Learning for Causal Decision Making

### 2019-03-03 UAI_2019/UAI_Draft.tex
cambios en la seccion de rationality

\section{Introduction}
Decision making under uncertain conditions is a fundamental task for intelligent reasoning and an any-day task for human beings and organizations (\cite{savage1954the}, \cite{danks2014unifying}, \cite{lake2017building}). It is a well known fact that if an agent has \textit{rational preferences} then he must choose according to the Maximum Expected Utility criteria, in which the agent chooses an action by considering both his preferences, which are represented by an utility function, and the probability of occurrence of outcomes: with respect to this point, the agent either knows the probabilities in the environment, or, if he does not, by considering a \textit{subjective} probability measure (\cite{von1944theory}, \cite{savage1954the}, \cite{bernardo2000bayesian}, \cite{gilboa2009decision}). The case where the uncertainty component in the decision problem is controlled by a \textit{causal mechanism}, has been previously studied in \cite{joyce1999foundations}, \cite{pearl2009causality}, \cite{lattimoreNIPS2016}, \cite{sen2017identifying}, \cite{gonzalez2018playing}, \cite{2019arXiv190202279G}. In particular, \cite{joyce1999foundations} provided the intuition that a decision maker should consider what is \textit{caused} by his actions, and \cite{pearl2009causality} proposed decision-making by maximizing a function involving the well known \textit{do} operator.

@@

From the human learning perspective, it has been shown that humans acquire and use causal information for decision-making and that humans even consider their actions as interventions in the world (\cite{hagmayer2009decision}, \cite{wellen2012learning} \cite{hagmayer2013repeated}). It has also been argued that humans tend to ignore probabilistic information and assign more importance to causal information (\cite{tversky1980causal}, \cite{pearl2009causality}). Human beings focus on \textit{local} aspects while learning causal relationswhich are later unified into a single structure (\cite{fernbach2009causal}, \cite{waldmann2008causal}, \cite{wellen2012learning}, \cite{danks2014unifying}). Following this idea, \cite{wellen2012learning} propose a model to explain how observations  and interventions are used by human beings to learn causal relations.

\section{Rationality, Decision Making and Expected Utility}
In this work we begin with a Decision Problem under Uncertainty where an agent, who has rational preferences, faces the problem of choosing one among many available actions with uncertain consequences. From the work of \cite{von1944theory} and \cite{savage1954the}, we know that the only criteria for making choices is the maximization of expected utility with respect to the agent's preferences and a probability distribution which represents either the probabilities of the uncertain events, or the agents own beliefs about such uncertainty. In \cite{gonzalez2018playing} and \cite{2019arXiv190202279G} the case where the uncertainty of the environment is controlled by a causal mechanism is considered. Since the (unknown) causal mechanism is itself a stochastic one, and the agent's preferences are rational, Savage's theorem state how an action must be chosen: maximizing the expected utility. If the decision maker does not know the causal mechanism, this is equivalent to not knowing the probabilities of the environment, so a subjective distribution must be used.

\section{Related Work}
In \cite{lattimoreNIPS2016}, the problem of finding a best intervention over a known causal model for a fixed number of learning trials is considered and it was found that by considering causal information a \textit{faster} optimal-action learning can be achieved. Their assumption of knowing the causal model was relaxed by \cite{sen2017identifying} who considers a partially known causal model and allows interventions over a the unknown part of the model. In \cite{gonzalez2018playing} it was proposed that a rational decision maker can find an optimal action by holding \textit{beliefs} about the existence and strength of causal relations and by using those beliefs \textit{as if} they were the true relations in order to make a choice and then updating those beliefs after observing what the chosen action \textit{caused} good actions can be found. Such procedure implicitly considers a distribution over Causal Graphical Models, where each structure is used within each step in otder to make a choice by considering his causal knowledge as it if were true, as proposed by \cite{joyce1999foundations}. Their work assumes that the decision maker knows the structure, but not the associated parameters, of the model. Later, \cite{2019arXiv190202279G}, such authors formalize that procedure by stating a formal criteria which has the form:
\[  a^\ast = \textrm{ argmax }_{a \in \mathcal{A}}\sum_g P(Y=1 | do(a), g)P(g), \]
where $\mathcal{A}$ is the set of available actions to the decision maker and $1$ is the most desired value of the \textit{objective variable} $Y$, for the decision maker. We notice that since the action chosen is the action with the highest probability of causally producing the most desired action, then $a^\ast$ necesarilly is the action that maximizes the expected utility for the agent. The function above described is indeed a counterfactual distribution, since it allows to compare the probability of causing a desired actions by alternative actions that were not chosen.

While \cite{pearl2009causality} proposes that a rational agent in the presence of a causal environment should optimize the function $U(x)$ defined as 
\[ U(x)=\sum_y P(y | do(x)) u(y), \]
we notice that this function does not consider the agent's beliefs, or current knowledge, about the causal structure that holds in his environment. Pearl's function assumes a single causal structure. as opposed by \cite{2019arXiv190202279G} who make use of an implicit probability distribution over causal structures.

>>>>

\section{Introduction}
Decision making under uncertain conditions is a fundamental task for intelligent reasoning and an any-day task for human beings and organizations (\cite{savage1954the}, \cite{danks2014unifying}, \cite{lake2017building}). It is a well known fact that if an agent has \textit{rational preferences} then he must choose according to the Maximum Expected Utility criteria, in which the agent chooses an action by considering both his preferences, which are represented by an utility function, and the probability of occurrence of outcomes: with respect to this point, the agent either knows the probabilities in the environment, or, if he does not, by considering a \textit{subjective} probability measure (\cite{von1944theory}, \cite{savage1954the}, \cite{bernardo2000bayesian}, \cite{gilboa2009decision}).

The case where the uncertainty component in the decision problem is controlled by a \textit{causal mechanism}, has been previously studied in \cite{joyce1999foundations}, \cite{pearl2009causality}, \cite{lattimoreNIPS2016}, \cite{sen2017identifying}, \cite{gonzalez2018playing}, \cite{2019arXiv190202279G}. In particular, \cite{joyce1999foundations} provided the intuition that a decision maker should consider what is \textit{caused} by his actions, and \cite{pearl2009causality} proposed decision-making by maximizing a function involving the well known \textit{do} operator.

@@

From the human learning perspective, it has been shown that humans acquire and use causal information for decision-making and that humans even consider their actions as interventions in the world (\cite{hagmayer2009decision}, \cite{wellen2012learning} \cite{hagmayer2013repeated}). It has also been argued that humans tend to ignore probabilistic information and assign more importance to causal information (\cite{tversky1980causal}, \cite{pearl2009causality}). Human beings focus on \textit{local} aspects while learning causal relationswhich are later unified into a single structure (\cite{fernbach2009causal}, \cite{waldmann2008causal},  \cite{danks2014unifying}). Following this idea, \cite{wellen2012learning} propose a model to explain how observations  and interventions are used by human beings to learn causal relations in terms of a local prediction-error learning. Following this line of thought, we propose here a local probabilistic encoding of the uncertainty that a decision maker has over the existence or not of causal relations between variables.

\section{Rationality and Decision Making under}
In this work we begin with a Decision Problem under Uncertainty where an agent, who has rational preferences, faces the problem of choosing one among many available actions with uncertain consequences. From the work of \cite{von1944theory} and \cite{savage1954the}, we know that the only criteria for making choices is the maximization of expected utility with respect to the agent's preferences and a probability distribution which represents either the probabilities of the uncertain events, or the agents own beliefs about such uncertainty.

Since \cite{joyce1999foundations} it has been argued that a decision maker should consider all of his beliefs about what will his decisions cause, while \cite{pearl2009causality} proposes that a rational agent in the presence of a causal environment should optimize the function $U(x)$ defined as 
\[ U(x)=\sum_y P(y | do(x)) u(y), \]
we notice that this function does not consider the agent's beliefs, or current knowledge, about the causal structure that holds in his environment as opposed to \cite{gonzalez2018playing} and \cite{2019arXiv190202279G} who make use of an implicit probability distribution over causal structures, while Pearl's function assumes a single causal structure.

\section{Related Work}

In \cite{lattimoreNIPS2016}, the problem of finding a best intervention over a known causal model for a fixed number of learning trials is considered and it was found that by considering causal information a \textit{faster} optimal-action learning can be achieved. Their assumption of knowing the causal model was relaxed by \cite{sen2017identifying} who considers a partially known causal model and allows interventions over a the unknown part of the model. In \cite{gonzalez2018playing} it was proposed that a rational decision maker can find an optimal action by holding \textit{beliefs} about the existence and strength of causal relations and by using those beliefs \textit{as if} they were the true relations in order to make a choice and then updating those beliefs after observing what the chosen action \textit{caused} good actions can be found. Such procedure implicitly considers a distribution over Causal Graphical Models, where each structure is used within each step in otder to make a choice by considering his causal knowledge as it if were true, as proposed by \cite{joyce1999foundations}. Their work assumes that the decision maker knows the structure, but not the associated parameters, of the model. Later, \cite{2019arXiv190202279G}, such authors formalize that procedure by stating a formal criteria which has the form:
\[  a^\ast = \textrm{ argmax }_{a \in \mathcal{A}}\sum_g P(Y=1 | do(a), g)P(g), \]
where $\mathcal{A}$ is the set of available actions to the decision maker and $1$ is the most desired value of the \textit{objective variable} $Y$, for the decision maker. We notice that since the action chosen is the action with the highest probability of causally producing the most desired action, then $a^\ast$ necesarilly is the action that maximizes the expected utility for the agent. The function above described is indeed a counterfactual distribution, since it allows to compare the probability of causing a desired actions by alternative actions that were not chosen.
